---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# worldcup26

An R wrapper for the [football-data.org v4 API](https://docs.football-data.org/general/v4/resources.html), focused on the 2026 FIFA Men's World Cup. Look up the participating teams, a team's full schedule, its next match, or its past results.

## Installation

```r
# install.packages("pak")
pak::pak("smach/worldcup26")
```

You need a free football-data.org API key. Register at
<https://www.football-data.org/client/register> and add the key to your `~/.Renviron`:

```
FOOTBALL_DATA_API_KEY=your-key-here
```

Restart R so the variable is picked up.

## Usage

```{r, eval = FALSE}
library(worldcup26)

list_teams()
team_schedule("USA")
team_next_match("Brazil")
team_past_results("Argentina")
```

Team lookups are flexible. All of these resolve to the same team:

```{r, eval = FALSE}
team_schedule("United States")
team_schedule("USA")
team_schedule("us")
```

Common aliases also work: `"Korea"` → South Korea, `"Czech Republic"` → Czechia, `"Cape Verde"` → Cape Verde Islands.

## Score display

Match scores are reported in a single `score_display` column:

| Situation                                    | Value                     |
|----------------------------------------------|---------------------------|
| Match yet to start                           | `""`                      |
| Match in progress                            | `"in progress"`           |
| Recently completed, no score posted yet      | `"no score available yet"`|
| Completed match with score                   | e.g. `"2–1"`         |
| Knockout decided on penalties                | e.g. `"1–1 (4–3 PK)"` |
| Postponed / cancelled / suspended            | `"postponed"`, etc.       |

The free football-data.org tier doesn't expose live in-match scores, so during a match you'll see `"in progress"`. Use a faster source if you need a live scoreline.

## Caching

API responses are cached to disk for one hour by default. Override with:

```r
options(worldcup26.cache_ttl = 600)  # 10 minutes
clear_cache()                         # nuke the cache manually
```

## Functions

* `list_teams()` — every team in the tournament
* `team_schedule(team)` — full schedule for a team
* `team_next_match(team)` — earliest upcoming match
* `team_past_results(team)` — already-played matches
* `all_matches()` — every World Cup match in one tibble
* `chat_data()` — the flat matches table used by the chat dashboard
* `worldcup26_chat()` — launches the natural-language chat dashboard
* `clear_cache()` — drop cached responses

## Natural-language chat dashboard

The package ships a small Shiny dashboard built on [querychat](https://github.com/posit-dev/querychat) that lets you ask questions in plain English:

```r
worldcup26::worldcup26_chat()
```

Sample questions:

- *When is Canada's next game?*
- *Show me all matches on June 15.*
- *Which teams are in Group D?*
- *List the round of 16 matches.*
- *Has Brazil played yet?*

The dashboard uses Anthropic Claude by default; you'll need an `ANTHROPIC_API_KEY` in your environment (`~/.Renviron` is a good place). Get a key at <https://console.anthropic.com>. To use a different provider or model, pass your own client:

```r
worldcup26_chat(client = ellmer::chat_openai(model = "gpt-4o"))
# or change the Claude model:
worldcup26_chat(model = "claude-opus-4-7")
```

## Companion Quarto site

This repo also ships a small Quarto + Observable JS site that lets visitors browse the schedule by team or by date. It builds on top of the package — an R chunk in `index.qmd` calls `list_teams()` and `all_matches()` at render time and hands the data to Observable JS via `ojs_define()`. The rendered page is fully static and needs no API key to view.

Site files (all excluded from the package build via `.Rbuildignore`):

* `index.qmd`, `_quarto.yml`, `_brand.yml`, `styles.css` — page source
* `.github/workflows/publish.yml` — rebuilds the page on every push to `main` and publishes to the `gh-pages` branch. A cron schedule rebuilds hourly during the match window (15:00–06:00 UTC / 11 AM–2 AM ET), skipping the overnight lull when no games are on
* `.github/workflows/R-CMD-check.yaml` — runs package checks on Ubuntu and Windows for pushes and pull requests

### Local preview

```bash
quarto preview
# or render once:
quarto render
```

You need the package installed (`R -e 'devtools::install()'` from the repo root) and `FOOTBALL_DATA_API_KEY` in your environment.

### GitHub Pages setup

1. Add a repository secret named `FOOTBALL_DATA_API_KEY` with your football-data.org API key (Settings → Secrets and variables → Actions).
2. Push to `main` or run the workflow manually from the Actions tab. It renders the site and pushes to `gh-pages`.
3. In Settings → Pages, choose **Branch: gh-pages** as the source.

The GitHub Actions workflow rebuilds the site hourly during the match window (15:00–06:00 UTC) and skips the quiet hours (07:00–14:00 UTC / 3–10 AM ET). Adjust the cron expression in `.github/workflows/publish.yml` if you want a different cadence. A prominent banner on the page tells visitors that live in-match scores are not available on the free API tier.

## License

MIT.
