---
title: "Getting started with worldcup26"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with worldcup26}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```

`worldcup26` wraps the football-data.org v4 API for the 2026 FIFA Men's World Cup. This vignette walks through the package's main functions end to end.

## Setup

Add your free API key to `~/.Renviron`:

```
FOOTBALL_DATA_API_KEY=your-key-here
```

Restart R, then load the package:

```{r}
library(worldcup26)
```

## List the participating teams

```{r}
teams <- list_teams()
teams
```

`teams` is a tibble with 48 rows (one per participating nation) and columns `id`, `name`, `short_name`, `tla` (three-letter abbreviation), and `crest_url`.

## Look up a team's schedule

`team_schedule()` accepts any reasonable variation on a team's name --- the full canonical name, a common short form, a three-letter code, or an entry from the built-in alias table:

```{r}
team_schedule("United States")
team_schedule("USA")
team_schedule("us")
```

All three return the same tibble. Times are shown in US Eastern (EDT), since every World Cup venue is in North America. Columns: `match_id`, `match_date` (a `Date` in US Eastern), `kickoff_edt` (e.g. `"9:00 PM EDT"`), `stage`, `group`, `home_team`, `away_team`, `status`, `score_display`, `venue`.

## What is the team's next match?

```{r}
team_next_match("Brazil")
```

Returns a single row (or an empty tibble if the team has no upcoming matches). "Upcoming" includes anything currently in progress.

## What about past results?

```{r}
team_past_results("Argentina")
```

If a match has finished and the API hasn't posted the score yet, `score_display` will read `"no score available yet"`. Live matches show `"in progress"`.

## All matches in one tibble

For aggregate work --- e.g. building a tournament tracker or feeding a Quarto dashboard --- pull every match in a single call:

```{r}
matches <- all_matches()
```

## Caching

API responses are cached on disk for one hour by default. To shorten the cache lifetime or clear it explicitly:

```{r}
options(worldcup26.cache_ttl = 600)  # 10 minutes
clear_cache()
```

## Natural-language chat dashboard

The package ships a small Shiny dashboard powered by [querychat](https://github.com/posit-dev/querychat). It lets you ask questions in plain English and the LLM translates them to SQL over a flat `matches` table:

```{r}
worldcup26_chat()
```

You need an `ANTHROPIC_API_KEY` in your environment. Get a key at <https://console.anthropic.com>. Sample questions:

- *When is Canada's next game?*
- *Show me all matches on June 15.*
- *Which teams are in Group D?*
- *List the round of 16 matches.*
- *Has Brazil played yet?*

The data the chat sees is whatever `chat_data()` returns. It's a denormalised tibble with one row per match plus convenience columns: `match_date`, `is_finished`, `is_upcoming`, `is_today`, three-letter codes, and a human-readable `stage_label`. Inspect it directly to understand what the chat is working with:

```{r}
chat_data()
```

## Reading the `score_display` column

| Situation                               | Value                                     |
|-----------------------------------------|-------------------------------------------|
| Match yet to start                      | `""`                                      |
| Match in progress                       | `"in progress"`                           |
| Recently completed, no score posted yet | `"no score available yet"`                |
| Completed match                         | e.g. `"2–1"`                         |
| Knockout decided on penalties           | e.g. `"1–1 (4–3 PK)"`           |
| Postponed / cancelled / suspended       | `"postponed"`, `"cancelled"`, `"suspended"` |

The free football-data.org tier does not surface live in-match scores. If you need real-time scoreline updates, pair `worldcup26` with another source for matches that are currently underway.
